An algorithm for finding the global extremum\\ of a partially defined function\thanks{This work was supported by the Russian Science Foundation, project No. 21-11-00204.}

### 2023-10-02 2023/MMST/samplepaper.tex
fix item 5 in modification discription
fix reviewer comments

Global optimization deals with the development of methods for solving problems of finding the global (absolute) minima of multidimensional multiextremal functions. The possibility of constructing a reliable estimate of the global extremum in such problems is principally based on the available information on the problem to be solved. This information makes it possible to relate the unknown values of the objective function to the known values at the points of the performed search trials.

@@

Using the Peano-Hilbert curves unambiguously mapping the interval $[0,1]$ onto the $N$-dimensional unit hypercube 
$$
D=\left\{ y \in R^N: -2^{-1} \leq y_i \leq 2^{-1}, 1 \leq i \leq N \right\} = \left\{ y(x): 0 \leq x \leq 1 \right\},
$$
we can reduce the initial problem (\ref{eq1}) to a one-dimensional problem
\begin{equation}\label{eq2} 
\phi^*=\phi(y(x^* ))=\min_{x \in [0,1]} \left\{ \phi(y(x)) \right\},
\end{equation}
that allows applying efficient one-dimensional optimization algorithms for its solving.

@@

Third, the objective function satisfies the Lipschitz condition
\begin{equation}\label{eq3} 
| \phi (y')-\phi (y'') | \leq L \| y'-y'' \|, \; y',y'' \in D,
\end{equation}
where $L$ is the Lipschitz constant. The dimensionality reduction scheme using Peano curves is known to juxtapose the multidimensional problem with a Lipschitz objective function (\ref{eq1}) to problem (\ref{eq2}) with the one-dimensional objective function satisfying H{\"o}lder condition
\begin{equation}\label{eq4} 
| \phi (x')-\phi (x'') | \leq K \| x'-x'' \| ^{1/N}, \; x',x'' \in [0,1],
\end{equation}
where $N$ is the dimensionality of the initial multidimensional problem and coefficient $K$ is related to the Lipschitz constant $L$ by the relation $K \leq 2L\sqrt {N+3}$ \cite{Strongin2000}.

@@

At every iteration of the global search a \textit{trial} is executed. We will refer to computing the objective function value  $\phi (y(x))$ from (\ref{eq2}) as a trial. The first trial will be executed in the middle point $x^1 \in (0,1)$. The choice of the point $x^{k+1}, k \leq 1$ of the next $(k+1)^\text{th}$ trial is performed on the base of the following rules.

\paragraph{Rule 1.} Renumber (by the lower indices) the points $x^i, 1 \leq i \leq k$  of the previous trials in the ascending order by their coordinates, i.e.
\begin{equation}\label{eq5} 
0=x_0 < x_1 < ... < x_i < ... < x_k=1
\end{equation}
and juxtapose them with the values $z_i=\phi (x_i), 1 \leq i \leq k$, computed at these points and to the indices $v_i=v(x_i)$ defined by the rule
\begin{equation}\label{eq6} 
v_i=v(x_i)=
  \begin{cases}
    -1 & {\quad \text{if } x_i \text{ is a boundary point,}}\\
    1  & {\quad \text{if } x_i \text{ is an internal point.}}
  \end{cases}
\end{equation}
The points $x_0=0$ and $x_{k+1}=1$ are introduced additionally (the values $z_0$ and $z_{k+1}$ are undefined) for convenience of the following notations.

@@

\paragraph{Rule 4.} For each interval $(x_{i-1},x_i),1 \leq i \leq k+1$, compute the value $R(i)$ called a \textit{characteristic} of the interval according to the formula
\begin{equation}\label{eq9} 
R(x_i)=
  \begin{cases}
    \Delta _i+\frac {{(z_i-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-1}-2z^*}{r \mu}, & {\quad  v(x_i)=v(x_{i-1})=1},\\
    2 \Delta _i-4 \frac {(z_i-z^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=1},\\
    2 \Delta _i-4 \frac {(z_{i-1}-z^*)}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=-1.}
  \end{cases}
\end{equation}

\paragraph{Rule 5.} Select the interval $(x_{t-1},x_t)$ with the largest value of characteristic $R(i)$:
\[
R(t)= \max\{R(i): \; 1 \leq i \leq k+1\}.
\]

@@

  \item For each interval $(x_{i-1}, x_i), 1 \leq i \leq k+1$, the computing of the characteristic $R(i)$ is performed according to the following formulae.

        \paragraph{Case 1.} For the intervals with \textbf{two computable points} or with \textbf{one computable and one boundary point}, apply the original rules of the global search algorithm
\begin{equation}\label{eq21} 
R(i)=
  \begin{cases}
     \Delta _i+\frac {{(z_i-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-1}-2z_1^*}{r \mu}, & {\quad  v(x_i)=v(x_{i-1})=1,}\\
    2 \Delta _i-4 \frac {(z_i-z_1^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=1,}\\
    2 \Delta _i-4 \frac {(z_{i-1}-z_1^*)}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=-1.}
  \end{cases}
\end{equation}
        \paragraph{Case 2.}   For the intervals with \textbf{one non-computable and one boundary point}, perform the imputation of the values for the non-computable point based on the neighbor computable point and apply the formula
\begin{equation}\label{eq22} 
R(i)=
  \begin{cases}
    2 \Delta _i-4 \frac {(z_{i+1}-z_0^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=0,}\\
    2 \Delta _i-4 \frac {(z_{i-2}-z_0^*)}{r \mu}, & {\quad  v(x_{i-1})=0, v(x_i)=-1.}
  \end{cases}
\end{equation}
        \paragraph{Case 3.} For the intervals with \textbf{one non-computable and one computable point}, perform the imputation of the value for the non-computable point based on the neighbor computable point and apply the formula
\begin{equation}\label{eq23} 
R(i)=
  \begin{cases}
     \Delta _i+\frac {{(z_i-z_{i-2})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-2}-2z_0^*}{r \mu}, & {\quad  v(x_{i-1})=0, v(x_i)=1,}\\
     \Delta _i+\frac {{(z_{i+1}-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_{i+1}+z_{i-1}-2z_0^*}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=0.}
  \end{cases}
\end{equation}
        \paragraph{Case 4.} For the intervals with \textbf{two non-computable points}, perform the imputation of the values for the non-computable points based on the rule (\ref{eq19}) and apply the formula
\begin{equation}\label{eq24} 
R(i)=\Delta _i+\frac {{(\tilde{z}_i-\tilde{z}_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {\tilde{z}_i+\tilde{z}_{i-1}-2z_1^*}{r \mu}, \; v(x_{i-1})=v(x_i)=0.
\end{equation}
        \paragraph{Case 5.} In the case when any imputation is impossible, use the formula
\begin{equation}\label{eq25} 
R(i)=\Delta _i \cdot {\left( 1-\frac {1}{r} \right)}^N+z_0^*.
\end{equation}
  \item If several intervals have equal characteristics, prefer the interval with computable boundary points that has the lowest number.
  \item If the best interval appears to be the one with two non-computable points, perform the next trial in the middle point of the interval.
\end{enumerate}

>>>>

Global optimization deals with the development of methods for solving problems of finding the global (absolute) minimum of multidimensional multiextremal functions. The possibility of constructing a reliable estimate of the global extremum in such problems is principally based on the available information on the problem to be solved. This information makes it possible to relate the unknown values of the objective function to the known values at the points of the performed search trials.

@@

Using the Peano-Hilbert curves unambiguously mapping the interval $[0,1]$ onto the $N$-dimensional unit hypercube 
$$
D=\left\{ y \in R^N: -2^{-1} \leq y_i \leq 2^{-1}, 1 \leq i \leq N \right\} = \left\{ y(x): 0 \leq x \leq 1 \right\},
$$
we can reduce the initial problem (\ref{eq1}) to a one-dimensional problem
\begin{equation}\label{eq2} 
f^*(x)=\phi(y(x^* ))=\min_{x \in [0,1]} \left\{ \phi(y(x)) \right\},
\end{equation}
that allows applying efficient one-dimensional optimization algorithms for its solving.

@@

Third, the objective function satisfies the Lipschitz condition
\begin{equation}\label{eq3} 
| \phi (y')-\phi (y'') | \leq L \| y'-y'' \|, \; y',y'' \in D,
\end{equation}
where $L$ is the Lipschitz constant. The dimensionality reduction scheme using Peano curves is known to juxtapose the multidimensional problem with a Lipschitz objective function (\ref{eq1}) to problem (\ref{eq2}) with the one-dimensional objective function satisfying H{\"o}lder condition
\begin{equation}\label{eq4} 
| \phi (x')-\phi (x'') | \leq K \rho(x',x''), \; x',x'' \in [0,1],
\end{equation}
where $\rho(x',x'') =  |x' - x''|^{1/N}$ and $N$ is the dimensionality of the initial multidimensional problem and coefficient $K$ is related to the Lipschitz constant $L$ by the relation $K \leq 2L\sqrt {N+3}$ \cite{Strongin2000}.

@@

At every iteration of the global search a \textit{trial} is executed. We will refer to computing the objective function value  $\phi (y(x))$ from (\ref{eq2}) as a trial. The first trial will be executed in the middle point $x^1 \in (0,1)$. The choice of the point $x^{k+1}, k \geq 1$ of the next $(k+1)^\text{th}$ trial is performed on the base of the following rules.

\paragraph{Rule 1.} Renumber (by the lower indices) the points $x^i, 0 \leq i \leq k$  of the previous trials in the ascending order by their coordinates, i.e.
\begin{equation}\label{eq5} 
0=x_0 < x_1 < ... < x_i < ... < x_{k}=1
\end{equation}
and juxtapose them with the values $z_i=\phi (x_i), 0 < i < k$, computed at these points and to the indices $v_i=v(x_i)$ defined by the rule
\begin{equation}\label{eq6} 
v_i=v(x_i)=
  \begin{cases}
    -1 & {\quad \text{if } x_i \text{ is a boundary point,}}\\
    1  & {\quad \text{if } x_i \text{ is an internal point.}}
  \end{cases}
\end{equation}
The points $x_0=0$ and $x_{k}=1$ are introduced additionally (the values $z_0$ and $z_{k}$ are undefined) for convenience of the following notations.

@@

\paragraph{Rule 4.} For each interval $(x_{i-1},x_i),1 \leq i \leq k$, compute the value $R(i)$ called a \textit{characteristic} of the interval according to the formula
\begin{equation}\label{eq9} 
R(x_i)=
  \begin{cases}
    \Delta _i+\frac {{(z_i-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-1}-2z^*}{r \mu}, & {\quad  v(x_i)=v(x_{i-1})=1},\\
    2 \Delta _i-4 \frac {(z_i-z^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=1},\\
    2 \Delta _i-4 \frac {(z_{i-1}-z^*)}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=-1.}
  \end{cases}
\end{equation}

\paragraph{Rule 5.} Select the interval $(x_{t-1},x_t)$ with the largest value of characteristic $R(i)$:
\[
R(t)= \max\{R(i): \; 1 \leq i \leq k\}.
\]

@@

  \item For each interval $(x_{i-1}, x_i), 1 \leq i \leq k$, the computing of the characteristic $R(i)$ is performed according to the following formulae.

        \paragraph{Case 1.} For the intervals with \textbf{two computable points} or with \textbf{one computable and one boundary point}, apply the original rules of the global search algorithm
\begin{equation}\label{eq21} 
R(i)=
  \begin{cases}
     \Delta _i+\frac {{(z_i-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-1}-2z_1^*}{r \mu}, & {\quad  v(x_i)=v(x_{i-1})=1,}\\
    2 \Delta _i-4 \frac {(z_i-z_1^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=1,}\\
    2 \Delta _i-4 \frac {(z_{i-1}-z_1^*)}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=-1.}
  \end{cases}
\end{equation}
        \paragraph{Case 2.}   For the intervals with \textbf{one non-computable and one boundary point}, perform the imputation of the values for the non-computable point based on the neighbor computable point and apply the formula
\begin{equation}\label{eq22} 
R(i)=
  \begin{cases}
    2 \Delta _i-4 \frac {(z_{i+1}-z_0^*)}{r \mu}, & {\quad  v(x_{i-1})=-1, v(x_i)=0,}\\
    2 \Delta _i-4 \frac {(z_{i-2}-z_0^*)}{r \mu}, & {\quad  v(x_{i-1})=0, v(x_i)=-1.}
  \end{cases}
\end{equation}
        \paragraph{Case 3.} For the intervals with \textbf{one non-computable and one computable point}, perform the imputation of the value for the non-computable point based on the neighbor computable point and apply the formula
\begin{equation}\label{eq23} 
R(i)=
  \begin{cases}
     \Delta _i+\frac {{(z_i-z_{i-2})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_i+z_{i-2}-2z_0^*}{r \mu}, & {\quad  v(x_{i-1})=0, v(x_i)=1,}\\
     \Delta _i+\frac {{(z_{i+1}-z_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {z_{i+1}+z_{i-1}-2z_0^*}{r \mu}, & {\quad  v(x_{i-1})=1, v(x_i)=0.}
  \end{cases}
\end{equation}
        \paragraph{Case 4.} For the intervals with \textbf{two non-computable points}, perform the imputation of the values for the non-computable points based on the rule (\ref{eq19}) and apply the formula
\begin{equation}\label{eq24} 
R(i)=\Delta _i+\frac {{(\tilde{z}_i-\tilde{z}_{i-1})}^2}{{(r \mu)}^2 \Delta _i} - 2 \frac {\tilde{z}_i+\tilde{z}_{i-1}-2z_1^*}{r \mu}, \; v(x_{i-1})=v(x_i)=0.
\end{equation}
        \paragraph{Case 5.} In the case when any imputation is impossible, use the formula
\begin{equation}\label{eq25} 
R(i)=\Delta _i \cdot {\left( 1-\frac {1}{r} \right)}^N+z_0^*.
\end{equation}
  \item If several intervals have equal characteristics, prefer the interval with computable points that has the lowest number.
  \item If the best interval appears to be the one with two non-computable points, perform the next trial in the middle point of the interval.
\end{enumerate}
